---
title: "Prac4"
output: html_document
date: "2024-10-31"
---
# Load the required libraries for raster and file handling
library(terra)
library(here)

# Load the raster layer for January temperature data
jan <- terra::rast("/Users/han/Desktop/CASA0005 /Prac3/prac3_data/wc2.1_5m_tavg_01.tif")

# Visualize the raster layer to examine its structure
plot(jan)

#Hey Han Guo, some comments that might help!
## Add short comments to describe what each step does.
## Remove Unused Libraries

# I have copied my response below - Hope this helps.

################
#Week 4 Homework
################


# Load the necessary libraries for spatial analysis, visualization, and data manipulation
library(sf)
library(tmap)
library(tmaptools)
library(RSQLite)
library(tidyverse)
library(here)
library(fs)
library(raster)
library(terra)
library(plotly)
library(dplyr)
library(stringr)
library(countrycode)
library(ggplot2)
library(readr)
library(janitor)
a.attr('aria-controls', id);                                                                               

# Load the Gender Inequality data
GenderInequality <- read_csv("HDR23-24_Composite_indices_complete_time_series.csv")

# Check the data type of each column
Datatypelist <- GenderInequality %>%
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to = "All_variables", 
               values_to = "Variable_class")

# Display the data type list
Datatypelist

# Load spatial data for the world (GeoJSON format)
world_map <- st_read(here::here("World_Countries_(Generalized)_9029012925078512962.geojson"))

# Extract the required columns for gender inequality data
gender_data <- GenderInequality[, c("iso3", "country", "gii_2010", "gii_2019")]

# Remove rows with missing values from gender_data
final_gender_data <- na.omit(gender_data)

# Create a new column showing the difference in gender inequality between 2010 and 2019
Inequality_Diff_between_2010_and_2019 <- final_gender_data %>%
  mutate(Inequalitydiff = gii_2010 - gii_2019)

# Convert ISO2 codes in world_map to ISO3 codes for consistency
world_map <- world_map %>%
  mutate(ISO3 = countrycode(ISO, "iso2c", "iso3c"))

# Merge spatial data with gender inequality data using ISO3 codes
world_data <- world_map %>%
  left_join(Inequality_Diff_between_2010_and_2019, by = c("ISO3" = "iso3"))

# Remove rows with missing values in the merged dataset
world_data <- na.omit(world_data)

# Visualize the change in gender inequality between 2010 and 2019 using a map
ggplot(data = world_data) +
  geom_sf(aes(fill = Inequalitydiff)) + # Color regions based on inequality difference
  scale_fill_viridis_c() + # Use a color scale suitable for continuous data
  theme_minimal() + # Apply a minimalistic theme
  labs(title = "Gender inequality changes between 2010 and 2019") # Add a descriptive title









```{r}
library(terra)
library(here)
jan<-terra::rast("/Users/han/Desktop/CASA0005 /Prac3/prac3_data/wc2.1_5m_tavg_01.tif")
# have a look at the raster layer jan
plot(jan)
```
```{r}
library(sf)
library(tmap) 
library(tmaptools)
library(RSQLite)
library(tidyverse)
library(here)
library(fs)
library(raster)
library(terra)
library(plotly)
library(dplyr)
library(stringr)
library(countrycode)

library(ggplot2)
library(readr)
library(janitor)
```
```{r}
#Load data Gender Inequality Data 
GenderInequality<-read_csv("HDR23-24_Composite_indices_complete_time_series.csv")
class(GenderInequality)

#Examing New Data
Datatypelist<-GenderInequality%>%
  summarise_all(class)%>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
  
```
```{r}
#Load spatial data of the world
library(sf)
world_map<-st_read(here::here("World_Countries_(Generalized)_9029012925078512962.geojson"))
```
```{r}
#extract the required gender inequality data
gender_data<-GenderInequality[, c("iso3","country","gii_2010","gii_2019")]
```


```{r}
#Use na.omit() to remove rows containing NA
final_gender_data<-na.omit(gender_data)

```
```{r}
#creat new column of difference in inequality between 2010 and 2019
Inequality_Diff_between_2010_and_2019<-final_gender_data %>%
  mutate(Inequalitydiff=gii_2010-gii_2019)
```

```{r}
#change to ISO3 in world_map
world_map<-world_map %>%
  mutate(ISO3 = countrycode(ISO, "iso2c", "iso3c"))

#Merge using country iso3
world_data<-world_map %>%
  left_join(Inequality_Diff_between_2010_and_2019, by = c("ISO3" = "iso3"))

#Use na.omit() to remove rows containing NA
world_data<-na.omit(world_data)
```

```{r}
#visulisation in map
library(ggplot2)

ggplot(data = world_data) +
  geom_sf(aes(fill = Inequalitydiff)) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Gender inequality changes between 2010 and 2019")
```

